### 2020-01-27 ChapterRampInput/ChapterRampInput.tex
Story 182: Intro complete

\section{Introduction  - Ramp input}

Measurements estimate the unknown value of a physical quantity, namely the measurand.
The to-be-measured physical quantity is applied as an input signal to a sensor. 
The sensor is a dynamical system and its output changes as a consequence of the input excitation and the sensor initial conditions.
The goal of a measurement is to estimate accurately the measurand value using the sensor transient response.
The transient response of a stable sensor decays to a steady state response.
In steady state, the most accurate estimation of the input true value is simply found using the sensor static gain.
However, the steady state is reached in theory after an infinite period of time and in practice we require fast estimations.
The trade-off between accuracy and speed exists in all measurements.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
\caption{ \label{fig:sensor_weight} The sensor transient response $y$ to an affine input excitation $u = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$. In the figure we observe an example of the input estimate $\widehat{u}$ obtained with the subspace method. The input parameters are calculated from $\widehat{u}$ using linear regression.}
\end{figure}

@@

\begin{table}[h!]
\centering
\caption{The values of the forgetting factor $\lambda$ and the samples shift $\tau$ that configure the subspace method for the different values of SNR. These values were obtained after calibration of the method and fixed during the simulation study.}
\begin{tabular}{|c c c c c c|} 
 \hline
 SNR \ [dB] & 20 & 30 & 40 & 50 & 60 \\ [0.5ex] 
 \hline
 $\lambda$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
 $\tau$ \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\label{table:lambdas}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
\caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimation decrease as the subspace method processes more samples. The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% after 400 and 500 samples, respectively ($T_s = 0.1$ ms). }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
\caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
\caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the interception estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
\caption{ \label{fig:CRLB_MSE_SNR_ab_dd_MC_10000} The Cram\'er-Rao lower bounds of the estimates $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_b$ determine the minimum uncertainty that can be achieved and increases with the measurement noise. 
The empirical mean squared errors $\mathrm{MSE}_{\hat{a}}$ and $\mathrm{MSE}_{\hat{b}}$ are near to the Cram\'er-Rao lower bounds within one order of magnitude for SNR smaller than 30 dB, and within two orders of magnitude for SNR between 40 dB and 60 dB. }
\end{figure}

@@

\begin{tabular}{|L|L L L L L|} 
 \hline
 \mathrm{Time \ [s]} & \mathrm{Mass \ [kg]} & 0.1 & 0.3 & 0.5 & 1.0 \\ [0.5ex] 
 \hline
 0.05 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.55\mathrm{x}10^{0} (@40 \ \mathrm{dB}) & 13.5\mathrm{x}10^{0} (@50 \ \mathrm{dB}) & 3.80\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.29\mathrm{x}10^{0}  (@50 \ \mathrm{dB}) \\ 
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-3} (@40 \ \mathrm{dB})  & 1.11\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.07\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 1.92\mathrm{x}10^{-2} (@60  \ \mathrm{dB})\\
 0.1 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.08\mathrm{x}10^{-2} (@30 \ \mathrm{dB}) & 1.16\mathrm{x}10^{0} (@40  \ \mathrm{dB}) & 8.34\mathrm{x}10^{-1} (@60 \ \mathrm{dB}) & 1.27\mathrm{x}10^{0} (@50  \ \mathrm{dB}) \\
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 5.99\mathrm{x}10^{-5} (@30 \ \mathrm{dB}) & 1.17\mathrm{x}10^{-1} (@50  \ \mathrm{dB}) & 2.10\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) & 3.76\mathrm{x}10^{-2} (@40  \ \mathrm{dB}) \\
 0.5 & \hspace{7.5mm} \mathrm{MSE}_{\hat{a}}: & 3.03\mathrm{x}10^{-2} (@20 \ \mathrm{dB}) & 2.98\mathrm{x}10^{-1} (@20  \ \mathrm{dB}) & 3.25\mathrm{x}10^{0} (@20 \ \mathrm{dB}) & 8.60\mathrm{x}10^{-2} (@20  \ \mathrm{dB}) \\
 & \hspace{7.5mm} \mathrm{MSE}_{\hat{b}}: & 3.10\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 9.86\mathrm{x}10^{-4} (@40 \ \mathrm{dB}) & 2.06\mathrm{x}10^{-5} (@50 \ \mathrm{dB}) & 1.74\mathrm{x}10^{-5} (@20 \ \mathrm{dB}) \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\
\label{table:differentmasses}
\end{table}

@@

\begin{tabular}{|C| C C C C|} 
\hline  
 &  & \sigma_{\mathrm{s}}: \hspace{4mm} 0\% & \hspace{8.1mm} 5\% & \hspace{8.1mm} 10\% \\ [0.5ex] 
 \hline
 \sigma_{{a},{b}}: \hspace{1.6mm} 0\% & \widehat{a} \ [\mathrm{kg/s}] & 0.990 \pm 9.6 \% \ (@50 \ \mathrm{dB}) &  0.999 \pm 8.2 \% \ (@20 \ \mathrm{dB}) & 0.996 \pm 8.2 \% \ (@50 \ \mathrm{dB}) \\ 
  & \widehat{b} \ \mathrm{[g]} & \hspace{-1.75mm} 10.0 \pm 10.3 \% \ (@20 \ \mathrm{dB}) & 9.91 \pm 14.2 \% \ (@30 \ \mathrm{dB}) &  9.99 \pm 21.7 \% \ (@60 \ \mathrm{dB}) \\ 
 \hspace{6.65mm} 5\% & \widehat{a} \ [\mathrm{kg/s}] & 0.994 \pm 9.7 \% \ (@50 \ \mathrm{dB}) & 0.993 \pm 11.3 \% \ (@20 \ \mathrm{dB}) & 0.997 \pm 8.0 \% \ (@50 \ \mathrm{dB}) \\  
  & \widehat{b} \ \mathrm{[g]} & 9.98 \pm 22.4 \% \ (@50 \ \mathrm{dB}) & \hspace{-1.75mm} 10.02 \pm 33.2 \% \ (@30 \ \mathrm{dB}) & 9.94 \pm 16.6 \% \ (@50 \ \mathrm{dB}) \\ 
\hspace{4.88mm} 10\% & \widehat{a} \ [\mathrm{kg/s}] & 1.00 \pm 20.3 \% \ (@40 \ \mathrm{dB}) & 1.00 \pm 22.0 \% \ (@50 \ \mathrm{dB}) & 0.996 \pm 17.7 \% \ (@40 \ \mathrm{dB}) \\    
 & \widehat{b} \ \mathrm{[g]} & 9.97 \pm 46.8 \% \ (@40 \ \mathrm{dB}) & 9.78 \pm 58.8 \% \ (@50 \ \mathrm{dB}) & 9.91 \pm 41.6 \% \ (@40 \ \mathrm{dB}) \\ [0.5ex] 
\hline
\end{tabular}
\
\label{table:dd_sensitivity}
\end{table}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
\caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor initial conditions are estimated with the ML method. After three iterations the relative errors of the estimates are smaller than 5 \%.  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
\caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. The variances of the affine input parameters estimates decrease faster than the variances of the initial conditions estimates.  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
\caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 
The relative errors of $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% only after 800 and 950 samples, respectively.
With the subspace method the relative errors are below 5 \% after 400 and 500 samples. }
\end{figure}

@@

A linear time-invariant weighing system is used as an test example for the estimation methods.
The weighing system becomes time-varying when an affine input excites the system.
The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
The subspace method results are also compared to those of an existing digital time-varying filter.
The coefficients of the time-varying filter require offline optimization.
The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cr\'amer-Rao Lower bound.

>>>>

\section{Introduction}

The goal of a measurement is to accurately estimate the input value using the sensor transient response.
The input might vary during the measurement being the simplest case the input that changes with a constant rate, described with an affine function.
The inputs that vary at a constant rate activate the sensor gradually. 
An affine input estimation method is needed to estimate the slope and the interception parameters of the affine function.

This chapter describes two signal processing methods for the estimation of affine input parameters.
These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local-optimization.
The subspace is a recursive method independent of the sensor model, and the ML method iteratively minimizes a cost function using a simulated sensor response in a receding time horizon.

In this chapter, the dynamic weighing is chosen as an implementation example.
The dynamic weighing of an affine input is an interesting metrology application because the affine input turns a linear time invariant sensor model into a time-varying sensor model.
The affine input estimation methods provide an insight on the properties of the dynamic measurement of time-varying inputs.

The effectiveness and the uncertainty of the methods are evaluated with simulation studies.
The uncertainty is assessed, for the subspace method, using a Taylor expansion of the exponentially weighted recursive least squares estimate, and
for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.

\begin{comment}
Measurements estimate the unknown value of a physical quantity, namely the measurand.
The to-be-measured physical quantity is applied as an input signal to a sensor. 
The sensor is a dynamical system and its output changes as a consequence of the input excitation and the sensor initial conditions.
The goal of a measurement is to estimate accurately the measurand value using the sensor transient response.
The transient response of a stable sensor decays to a steady state response.
In steady state, the most accurate estimation of the input true value is simply found using the sensor static gain.
However, the steady state is reached in theory after an infinite period of time and in practice we require fast estimations.
The trade-off between accuracy and speed exists in all measurements.

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2.pdf} 
\caption{ \label{fig:sensor_weight} The sensor transient response $y$ to an affine input excitation $u = at+b$ is processed by the estimation methods to estimate the parameters $a$ and $b$. In the figure we observe an example of the input estimate $\widehat{u}$ obtained with the subspace method. The input parameters are calculated from $\widehat{u}$ using linear regression.}
\end{figure}

@@

\begin{table}[h!]
\centering
\caption{The values of the forgetting factor $\lambda$ and the samples shift $\tau$ that configure the subspace method for the different values of SNR. These values were obtained after calibration of the method and fixed during the simulation study.}
\begin{tabular}{c| c c c c c} 
 \hline
 SNR \ [dB] & 20 & 30 & 40 & 50 & 60 \\ [0.5ex] 
 \hline
 $\lambda$ & 0.939 & 0.940 & 0.955 & 0.959 & 0.959 \\ % forgetting factor
 $\tau$ \  & 15 & 15 & 17 & 14 & 20 \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\label{table:lambdas}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_3.pdf} 
\caption{ \label{fig:rele_dd_40dB_s1} The relative errors of the affine input parameters estimation decrease as the subspace method processes more samples. The relative errors of the estimates $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% after 400 and 500 samples, respectively ($T_s = 0.1$ ms). }
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_4.pdf} 
\caption{ \label{fig:rele_SNR_dd_10000} The minimum value of the estimation relative errors obtained with the subspace method is less than 2  \% regardless of the SNR between 20 dB and 60 dB. }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_5.pdf} 
\caption{ \label{fig:CRLB_MSE_ab_dd_40dB_MC_10000} When the SNR of the sensor response is 40 dB, the mean squared errors of the slope estimate $\widehat{a}$ and the interception estimate $\widehat{b}$, obtained by the subspace method, are two orders of magnitude above the theoretical minimum variance given by the Cram\'er-Rao lower bound.}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_6.pdf} 
\caption{ \label{fig:CRLB_MSE_SNR_ab_dd_MC_10000} The Cram\'er-Rao lower bounds of the estimates $\mathrm{CRLB}_a$ and $\mathrm{CRLB}_b$ determine the minimum uncertainty that can be achieved and increases with the measurement noise. 
The empirical mean squared errors $\mathrm{MSE}_{\hat{a}}$ and $\mathrm{MSE}_{\hat{b}}$ are near to the Cram\'er-Rao lower bounds within one order of magnitude for SNR smaller than 30 dB, and within two orders of magnitude for SNR between 40 dB and 60 dB. }
\end{figure}

@@

\begin{tabular}{L|L L L L} 
 \hline
 & \mathrm{Mass} & \\ [0.5ex] 
 \mathrm{Time} & 0.1 \ kg & 0.3 \ kg & 0.5 \ kg & 1.0 \ kg \\ 
 \hline
\multicolumn{1}{l|}{0.05 \ s} & \\
 \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.5\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 13.5\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} & 3.8\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\ 
\multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{b}} \ [g^2]} &  3.1\mathrm{x}10^{-3}_{(@40 \ \mathrm{dB})}  & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 1.9\mathrm{x}10^{-2}_{(@60 \ \mathrm{dB})} \\
\multicolumn{1}{l|}{0.1 \ s} & \\
 \multicolumn{1}{r|}{\hspace{1mm}\mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.1\mathrm{x}10^{-2}_{(@30 \ \mathrm{dB})} & 1.1\mathrm{x}10^{0}_{(@40 \ \mathrm{dB})} & 8.3\mathrm{x}10^{-1}_{(@60 \ \mathrm{dB})} & 1.3\mathrm{x}10^{0}_{(@50 \ \mathrm{dB})} \\
\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 6.0\mathrm{x}10^{-5}_{(@30 \ \mathrm{dB})} & 1.2\mathrm{x}10^{-1}_{(@50 \ \mathrm{dB})} & 2.1\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} & 3.7\mathrm{x}10^{-2}_{(@40 \ \mathrm{dB})} \\
\multicolumn{1}{l|}{0.5 \ s} & \\
 \multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{a}} \ [g^2]} & 3.0\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} & 3.0\mathrm{x}10^{-1}_{(@20 \ \mathrm{dB})} & 3.2\mathrm{x}10^{0}_{(@20 \ \mathrm{dB})} & 8.6\mathrm{x}10^{-2}_{(@20 \ \mathrm{dB})} \\
\multicolumn{1}{r|}{\hspace{1mm} \mathrm{MSE}_{\hat{b}} \ [g^2]} & 3.1\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 9.8\mathrm{x}10^{-4}_{(@40 \ \mathrm{dB})} & 2.0\mathrm{x}10^{-5}_{(@50 \ \mathrm{dB})} & 1.7\mathrm{x}10^{-5}_{(@20 \ \mathrm{dB})} \\ [0.5ex] % [\mathrm{samples}]
 \hline
\end{tabular}
\
\label{table:differentmasses}
\end{table}

@@

\begin{tabular}{C| C C C} 
\hline  
& \sigma_{\mathrm{s}} \\ [0.5ex] 
\sigma_{{a},{b}}  & 0\% & 5\% & 10\% \\  
 \hline
 \multicolumn{1}{l|}{0\%} \\
 \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.6 \%_{(@50 \ \mathrm{dB})} &  1.0 \pm 8.2 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.2 \%_{(@50 \ \mathrm{dB})} \\ 
 \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 10.3 \%_{(@20 \ \mathrm{dB})} & 9.9 \pm 14.2 \%_{(@30 \ \mathrm{dB})} &  10.0 \pm 21.7 \%_{(@60 \ \mathrm{dB})} \\ 
 \multicolumn{1}{l|}{5\%} \\
 \multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 9.7 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 11.3 \%_{(@20 \ \mathrm{dB})} & 1.0 \pm 8.0 \%_{(@50 \ \mathrm{dB})} \\  
 \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 10.0 \pm 22.4 \%_{(@50 \ \mathrm{dB})} & 10.0 \pm 33.2 \%_{(@30 \ \mathrm{dB})} & 9.9 \pm 16.6 \%_{(@50 \ \mathrm{dB})} \\ 
\multicolumn{1}{l|}{10\%} \\
\multicolumn{1}{l|}{\hspace{2mm} \widehat{a} \ [\mathrm{kg/s}]} & 1.0 \pm 20.3 \%_{(@40 \ \mathrm{dB})} & 1.0 \pm 22.0 \%_{(@50 \ \mathrm{dB})} & 1.0 \pm 17.7 \%_{(@40 \ \mathrm{dB})} \\    
 \multicolumn{1}{l|}{\hspace{2mm} \widehat{b} \ \mathrm{[g]}} & 9.9 \pm 46.8 \%_{(@40 \ \mathrm{dB})} & 9.8 \pm 58.8 \%_{(@50 \ \mathrm{dB})} & 9.9 \pm 41.6 \%_{(@40 \ \mathrm{dB})} \\ [0.5ex] 
\hline
\end{tabular}
\
\label{table:dd_sensitivity}
\end{table}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_7.pdf} 
\caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor initial conditions are estimated with the ML method. After three iterations the relative errors of the estimates are smaller than 5 \%.  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_8.pdf} 
\caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. The variances of the affine input parameters estimates decrease faster than the variances of the initial conditions estimates.  }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_9.pdf} 
\caption{ \label{fig:rele_tv_40dB_s1} The relative errors of the time-varying filter estimation converge slower than with the subspace method. 
The relative errors of $\widehat{a}$ and $\widehat{b}$ are smaller than 5 \% only after 800 and 950 samples, respectively.
With the subspace method the relative errors are below 5 \% after 400 and 500 samples. }
\end{figure}

@@

A linear time-invariant weighing system is used as an test example for the estimation methods.
The weighing system becomes time-varying when an affine input excites the system.
The estimation methods are compared in a simulation study where the time-varying sensor response is perturbed by measurement noise, that is assumed white of zero mean and known variance.
The subspace method results are also compared to those of an existing digital time-varying filter.
The coefficients of the time-varying filter require offline optimization.
The estimation results obtained with the subspace method converges two times faster and is one order of magnitude smaller than those obtained with the time-varying filter.
The empirical mean squared errors of the subspace method estimation is two orders of magnitude larger than the theoretical minimum given by the Cram\'er-Rao Lower bound.
